Learning Auditory Similarities Between Instrumental Playing Techniques

### 2020-05-14 paper/paper.tex
recall the definition of S2

Moreover, we evaluate the nearest neighbor algorithm $\mathbf{\Phi}$ in the absence of any metric learning at all.
This corresponds to using a Euclidean distance to compare scattering coefficients, i.e., to set $\mathbf{L}$ to the identity matrix.
Note that the runtime complexity of Euclidean nearest-neighbor search is identical to LMNN search.
Over our cohort of $K=31$ subjects, we reported a precision at rank five (P@5) of $\textbf{92.9\%} \pm 3$, which is noticeably worse than the best performing system.
This gap in performance demonstrates the importance of complementing unsupervised feature extraction by supervised metric learning in the design  of computational models for timbre similarity between instrumental playing techniques.

@@

Our best performing system operates with joint time--frequency scattering coefficients as spectrotemporal modulation features.
These features are extracted within a temporal context of duration equal to $T=\SI{1000}{\milli\second}$.
This value is considerably larger than the frame size of purely spectral features, such as spectral centroid, spectral flux, or mel-frequency cepstral coefficients (MFCC).

When considering the commonly used duration of $25~\mathrm{ms}$, the performance drops down to $\textbf{90.9\%} \pm 4$. This demonstrates that the joint scattering is able to benefit from a larger time support, possibly to capture longer-term modulations which play an important part when considering extended playing techniques \cite{lostanlen2018extended}.

@@

Considering the separable scattering gives a performance of $91\% \pm 4.5$. \ml{ vl. explain separable vs joint. write details on what is lost here ?}
% TODO VL: write about frequency-depedendent time shift. cite IEEE TSP

Further simplifying the feature set by replacing the scattering coefficients with MFCCs gives a performance of $81.8\% \pm 7.2$.
Note that the MFCCs are equivalent to first-order separable scattering coefficients \cite{anden2014tsp}.
Compared to the separable scattering (which includes both first- and second-order coefficients), we see that the MFCCs perform markedly worse.
In order to confirm that the additional information captured by the joint scattering is effectively important, another feature set is considered where the MFCCs are complemented with monomials in order to reach the dimensionality of the joint scattering.
\ja{Explain why this is necessary.}%
To do so, the $40$ features $f_i$ of the MFCCs are multiplied together $f_{i, j} = f_i \times f_j, i \neq j$.
\ja{What are the $f_i$ here?
How are they defined?}%
% TODO VL: give reference to MFCCs.
The resulting feature set consists of the 40 MFCCs features $f_i$ and the $1140$ features $f_{i, j}$.
The resulting $ap@5$ of $81.5\% \pm 7.3$ indicates that the increase of dimensionality is has a small effect, confirming the relevance of the additional information provided by the joint scattering.
\ml{vl outro on modulations}
% TODO VL: insist on the role of modulations in playing techniques.

>>>>

Moreover, we evaluate the nearest neighbor algorithm $\mathbf{\Phi}$ in the absence of any metric learning at all.
This corresponds to using a Euclidean distance to compare scattering coefficients, i.e., to set $\mathbf{L}$ to the identity matrix.
Note that the runtime complexity of Euclidean nearest-neighbor search is identical to LMNN search.
Over our cohort of $K=31$ subjects, we report a precision at rank five (P@5) of $\textbf{92.9\%} \pm 3$, which is noticeably worse than the best performing system.
This gap in performance demonstrates the importance of complementing unsupervised feature extraction by supervised metric learning in the design  of computational models for timbre similarity between instrumental playing techniques.

@@

Our best performing system operates with joint time--frequency scattering coefficients as spectrotemporal modulation features.
These features are extracted within a temporal context of duration equal to $T=\SI{1000}{\milli\second}$.
This value is considerably larger than the frame size of purely spectral features, such as spectral centroid, spectral flux, or mel-frequency cepstral coefficients (MFCC).
Indeed, the frame size of spectral features for machine listening is typically set to $T=\SI{23}{\milli\second}$, i.e., $2^{10}=1024$ samples at a sampling rate of $\SI{44,1}{\kilo\hertz}$.

As a point of comparison, we set the maximum time scale of joint time--frequency scattering coefficients to $T=\SI{25}{\milli\second}$, hence a $40$-fold reduction in context size.
Over our cohort of $K=31$ subjects, we report a precision at rank five (P@5) of $\textbf{90.9\%} \pm 4$, which is noticeably worse than the best performing system.
This gap in performance extends the findings of a previous publication \cite{lostanlen2018extended}, which reported that metric learning with temporal scattering coefficients tends to improve with growing values of $T$, until reaching a plateau of performance around $T\sim\SI{500}{\milli\second}$.

@@

Lastly, we consider the impact of alternative choices for feature extraction.
Let us recall the full definition of joint time--frequency scattering:
\begin{equation}
\mathbf{S_2}\boldsymbol{x}(t,\lambda,\alpha,\beta) = 
\Big\vert
\big\vert
\boldsymbol{x} \overset{t}{\ast} \boldsymbol{\psi}_{\lambda}
\big\vert
\overset{t}{\ast}
\boldsymbol{\psi}_{\alpha}
\overset{\log_2 \lambda}{\ast}
\boldsymbol{\psi}_{\beta}
\Big\vert
\overset{t}{\ast} \boldsymbol{\phi}_{T}
\overset{\log \lambda}{\ast} \boldsymbol{\phi}_{F},
\end{equation}
where the symbol $\boldsymbol{\psi}_{\lambda}$ (resp. $\boldsymbol{\phi}_T$) denotes a Morlet wavelet of center frequency $\lambda$ (resp. a Gaussian low-pass filter of scale $T$).

Considering the separable scattering gives a performance of $91\% \pm 4.5$.
% TODO VL: write about frequency-depedendent time shift. cite IEEE TSP

Further simplifying the feature set by replacing the scattering coefficients with MFCCs gives a performance of $81.8\% \pm 7.2$.
Compared to the separable scattering (which includes both first- and second-order coefficients), we see that the MFCCs perform markedly worse.
In order to confirm that the additional information captured by the joint scattering is effectively important, another feature set is considered where the MFCCs are complemented with monomials in order to reach the dimensionality of the joint scattering.
To do so, the $40$ features $f_i$ of the MFCCs are multiplied together $f_{i, j} = f_i \times f_j, i \neq j$.
\ja{What are the $f_i$ here?
How are they defined?}%
% TODO VL: give reference to MFCCs.
The resulting feature set consists of the 40 MFCCs features $f_i$ and the $1140$ features $f_{i, j}$.
The resulting $ap@5$ of $81.5\% \pm 7.3$ indicates that the increase of dimensionality is has a small effect, confirming the relevance of the additional information provided by the joint scattering.
% TODO VL: insist on the role of modulations in playing techniques.
